---
title: "STAT482-Time series"
author: "Huiyan Sang"
date: 
output: slidy_presentation
--- 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.height="60%", out.width="60%")
```

## Time Series
- A time series is simply a series of data points ordered in time. 
- The goal is to study the temporal pattern and/or make a forecast for the future.
- Examples:
  + Stock price
  + Electricity consumption
  + Daily temperature 
  
## Time series

- Is the times series autocorrelated?
- Is it stationary?
- Is there a seasonality?

## R packages
- xts
- zoo
- forecast
- quantmod
- tidyquant

## Summary with time series object

\tiny
```{r, echo = TRUE}
library(forecast)
library(ggplot2)
head(gas)
start(gas)
end(gas)
frequency(gas)
```
## Plot with time series object
\tiny
```{r,echo=TRUE,out.height="70%", out.width="70%"}
autoplot(gas) ## this is a time series object
tseries_sub <- window(gas, start=c(1990, 1), end=c(1994,12))
print(tseries_sub)
is.ts(gas)
```
\normalsize
## Convert your data to a time series object

\tiny
```{r,echo=TRUE,out.height="70%", out.width="70%"}
head(economics,2)
tail(economics ,2)
```

## Convert your data to a time series object and plot it
```{r,echo=TRUE}
dat_ts <- ts(economics$unemploy, start = c(1967, 7), end = c(2015, 4), frequency = 12) 
autoplot(dat_ts)
```

## Convert your data to a multivariate time series object and plot it

```{r,echo=TRUE}
dat_ts <- ts(economics[,2:6], start = c(1967, 7), end = c(2015, 4), frequency = 12) 
autoplot(dat_ts[,c('pop','uempmed','unemploy')]) ## plot multiple time series 
```


  
## Autocorrelations
- Autocorrelation, also called serial correlation, is the correlation of a signal with a delayed copy of itself as a function of delay. 
- It measures the similarity between observations as a function of the time lag between them. 
  + Lag-1 autocorrelation, $\rho(1)=\text{corr}(x_t,x_{t-1})$
  + Lag-2 autocorrelation, $\rho(2)=\text{corr}(x_t,x_{t-2})$
  + $\rho(k)$ is called the autocorrelation function that shows lag-k correlations 

## Autocorrelations 

```{r, echo = TRUE} 
ggAcf(gas)
```
 
## More on autocorrelations
```{r, echo = TRUE} 
g1=ggAcf(nhtemp) ## Annual temperatures in New Haven
g2=ggAcf(nottem) ## Monthly temperatures at Nottingham
gridExtra::grid.arrange(g1,g2,nrow=2)
```

## Seasonality

Additive decompositions
Time series = Seasonal + Trend + Random 

```{r, echo = TRUE}  
decom.ts=decompose(nottem)
plot(decom.ts)
```
 

## Fit an autoregressive integrated moving averages model. 

auto.arima() automatically determines orders and check for seasonality etc. 

```{r,echo=TRUE}
fit <- auto.arima(gas) 
#accuracy(fit)
```


## Forecasting usingthe fitted arima model
\tiny
```{r,echo=TRUE}
forecast(fit, 1) 
plot(forecast(fit, 1)) 
```
 

## Stationarity vs Nonstationarity
- Definition
If $\{y_t\}$ is a stationary time series, then for all $s$, the
distribution of $(y_t,\ldots, y_{t+s})$ does not depend on $t$. 
- Its statistical properties (mean, variance and covariance among time series) do not change over time.
- How to check stationarity?
- How to convert nonstationary TS to stationary TS?

## Stationarity vs Nonstationarity

```{r,echo=TRUE}
library(tseries) 
plot(gas) 
```
 
## Augmented Dickey-Fuller test
\tiny
```{r,echo=TRUE} 
adf.test(gas)  
```
 

## Take differences to make the time series stationary
\tiny
```{r,echo=TRUE}
adf.test(diff(gas))
```

## Relationship between two time series
\tiny
Cross-correlation function (CCF) is defined analogously to the ACF. It measures the lagged correlation between two time series. $\rho(k)=\text{corr}(x_t,y_{t-k})$

```{r,echo=TRUE}  
dat_ts <- ts(economics[,c('pop','unemploy')], 
             start = c(1967, 7), end = c(2015, 4), 
             frequency = 12) 
pop <- diff (dat_ts[,1])
unemploy <- diff(dat_ts[,2])
```

## CCF plot

```{r,echo=TRUE}
ccf (pop[-c(1:7)], unemploy[-c(1:7)])
```
